---
title: "Alcohol Consumption and Happiness Code"
subtitle: "Soc 225: Data & Society"
author: "Lab 4: Jaclyn Brockschmidt, Devin English, Maelisha Lee, Stassney Obregon, and Jennifer Qin"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
---
# Data Collection

# Data Preparation
## Install packages
```{r}
#install.packages("tidyverse")
#install.packages("gridExtra")
```

## Load libraries
```{r}
library(tidyverse)
library(rvest)
library(httr)
library(tidyverse)
library(gridExtra)
```

# Data preparation

## Load CSV data
```{r}
worldHapp <- read_csv("data/WorldHappiness-2016.csv", na = c("NA", "N/A"))
alcCons <- read_csv("data/HappinessAlcoholConsumption-2016.csv", na = c("NA", "N/A"))
```

## Scrape minimum drinking ages
```{r}
procon_page <- 
  read_html("https://drinkingage.procon.org/minimum-legal-drinking-age-in-other-countries/")
drinking_table <- html_node(procon_page, 
                     css = "table.newblue-table-2")
minDrinkAges <- html_table(drinking_table) %>%
  as_tibble()
#minDrinkAges <- as_tibble(minDrinkAges)
colnames(minDrinkAges) <- c("Country", "OnPremisePurchaseAge", "OffPremisePurchaseAge", "Notes")
minDrinkAges <- minDrinkAges %>% filter(Country != "Country")

minDrinkAges
```

## Merge data
```{r}
# Perform an inner join by country
alcKeep = c("Country", "Hemisphere", "HDI", "GDP_PerCapita", "Beer_PerCapita", "Spirit_PerCapita", "Wine_PerCapita")
worldAndAlc <- merge(worldHapp, alcCons[, alcKeep], by="Country", all=FALSE)
allData <- merge(worldAndAlc, minDrinkAges, by="Country", all=FALSE)

allData
```

## Row loss statistics
```{r}
rowsBefore <- max(nrow(worldHapp), nrow(alcCons), nrow(minDrinkAges))
rowsAfter <- nrow(allData)
rowLoss <- rowsBefore - rowsAfter
sprintf("%d/%d rows kept (%d countries excluded)", rowsAfter, rowsBefore, rowLoss)
```

## Write data
```{r}
write.csv(minDrinkAges, "data/MinimumDrinkingAges-2016.csv", row.names=FALSE)
write.csv(allData, "data/HappinessAlcoholRegulations-2016.csv", row.names=FALSE)
```

## Read Data
```{r}
alcData <- read_csv("data/HappinessAlcoholRegulations-2016.csv")
```

## Create Total Alcohol Consumption Per Capita
```{r}
alcData <- alcData %>% mutate(alcTotal = (Spirit_PerCapita + Wine_PerCapita + Beer_PerCapita))
```

## Create Separate Dataframe For Countries Without Age Restrictions
```{r}
alcDataNoAge <- alcData %>%
  filter(OnPremisePurchaseAge == "No age minimum")
```


## Recoding Purchase Age Data to Numeric
```{r}
alcData <- alcData %>%
  mutate(OnPremisePurchaseAge = parse_number(OnPremisePurchaseAge))
```

## Remove NA
```{r}
rowsBefore <- nrow(alcData)
alcData <- alcData[!is.na(alcData$OnPremisePurchaseAge), ]
rowsAfter <- nrow(alcData)
sprintf("%d/%d rows removed", rowsBefore - rowsAfter, rowsBefore, rowsAfter)

alcData
```

# Plots
## How Alcohol Consumption Relates and \nLegal Drinking Age to Happiness Scores by Country
```{r}
alcData <- alcData %>% arrange(alcData$"Happiness Rank")
top <- alcData %>% head(5)
bottom <- alcData %>% tail(5)
alc<- full_join(top,bottom)

plot <- alc %>%
  ggplot(aes(x = factor(Country,c('Denmark', 'Switzerland', 'Iceland', 'Norway', 'Finland', 'Chad', 'Madagascar', 'Tanzania', 'Liberia', 'Syria')), 
             y = `alcTotal`, fill = factor(OnPremisePurchaseAge))) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("Top 5 and Bottom 5 Happiness Scoring Countries (Left to Right)") +
  ylab("Alcohol Consumption Per Capita") +
  labs(title = "How Alcohol Consumption Relates and \nLegal Drinking Age to Happiness Scores by Country", fill='Legal Drinking Age')
plot + scale_fill_brewer(palette="Blues")
```

## How Alcohol Consumption and GDP Relates to Happiness Scores by Country
```{r}
alcData <- alcData %>% arrange(alcData$"Happiness Rank")
top <- alcData %>% head(5)
bottom <- alcData %>% tail(5)
alc<- full_join(top,bottom)

alc %>%
  ggplot(aes(x = factor(Country,c('Denmark', 'Switzerland', 'Iceland', 'Norway', 'Finland', 'Chad', 'Madagascar', 'Tanzania', 'Liberia', 'Syria')), 
             y = `alcTotal`, fill = GDP_PerCapita)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  xlab("Top 5 and Bottom 5 Happiness Scoring Countries (Left to Right)") +
  ylab("Alcohol Consumption Per Capita") +
  labs(title = "How Alcohol Consumption and GDP Relates to Happiness Scores by Country", fill='GDP Per Capita')
```

## How Alcohol Consumption Relates to Legal Drinking Age by Region
```{r}
byRegion <- alcData

a <- byRegion %>% filter(Region == "Australia and New Zealand") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
a <- a %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(15,16,17,18,19,20,21)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Australia and \nNew Zealand")

ce <- byRegion %>% filter(Region == "Central and Eastern Europe") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
ce <- ce %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(18,20)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Central and \nEastern Europe")

ea <- byRegion %>% filter(Region == "Eastern Asia") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
ea <- ea %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(15,16,17,18,19,20,21)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Eastern Asia")

la <- byRegion %>% filter(Region == "Latin America and Caribbean") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
la <- la %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(16,18,20)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Latin America and \nCaribbean")

me <- byRegion %>% filter(Region == "Middle East and Northern Africa") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
me <- me %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(16,18)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Middle East and \nNorthern Africa")

na <- byRegion %>% filter(Region == "North America") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
na <- na %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(18,21)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "North America")

sea <- byRegion %>% filter(Region == "Southeastern Asia") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
sea <- sea %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(18,20)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "SouthEastern Asia")

sa <- byRegion %>% filter(Region == "Sub-Saharan Africa") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
sa <- sa %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(15,16,17,18,19,20,21)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +  ylim(0, 700) +
  labs(title = "Sub-Saharan Africa")

we <- byRegion %>% filter(Region == "Western Europe") %>% group_by(OnPremisePurchaseAge) %>% summarise(mean = mean(alcTotal))
we <- we %>%
  ggplot(aes(x = `OnPremisePurchaseAge`, 
             y = mean)) +
  geom_col(fill = "blue3") +
  scale_x_continuous(breaks=c(15,16,17,18,19,20,21)) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),axis.title.x=element_blank(), axis.title.y=element_blank(), plot.title = element_text(size=10, hjust = 0.5)) +
  ylim(0, 700) +
  labs(title = "Western Europe")

grid.arrange(a, ce, ea, la, me, na, sea, sa, we, ncol=3, top="How Alcohol Consumption Relates to Legal Drinking Age by Region",
             bottom="Legal Drinking Age", 
               left="Alcohol Consumption Per Capita")
```

## How Happiness Differs in Countries With and Without Age Restrictions
```{r}
meanWithAge <- mean(alcData$"Happiness Score")
minWithAge <- min(alcData$"Happiness Score")
maxWithAge <- max(alcData$"Happiness Score")

meanNoAge <- mean(alcDataNoAge$"Happiness Score")
minNoAge <- min(alcDataNoAge$"Happiness Score")
maxNoAge <- max(alcDataNoAge$"Happiness Score")

stat <- c(meanWithAge, meanNoAge, minWithAge, minNoAge, maxWithAge, maxNoAge)
stat_type <- c("mean", "mean", "min", "min", "max", "max")
stat_label <- c("Age Restrictions", "No Age Restrictions", "Age Restrictions", "No Age Restrictions", "Age Restrictions", "No Age Restrictions")
withWithoutAge <- data.frame(stat, stat_type, stat_label)

withWithoutAge %>%
  ggplot(aes(x = stat_label,
             y = stat,
             fill = factor(stat_label))) +
  geom_col() +
  facet_wrap(~stat_type) +
  xlab("Summary Statistics") +
  ylab("Happiness Score") +
  labs(title = "How Happiness Differs In Countries\nWith and Without Drinking Age Restrictions", fill='Drinking Restrictions') +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_fill_brewer(palette = "Paired")
```

## How Alcohol Consumption Differs in Countries With and Without Age Restrictions
```{r}
meanWithAge <- mean(alcData$alcTotal)
minWithAge <- min(alcData$alcTotal)
maxWithAge <- max(alcData$alcTotal)

meanNoAge <- mean(alcDataNoAge$alcTotal)
minNoAge <- min(alcDataNoAge$alcTotal)
maxNoAge <- max(alcDataNoAge$alcTotal)

stat <- c(meanWithAge, meanNoAge, minWithAge, minNoAge, maxWithAge, maxNoAge)
stat_type <- c("mean", "mean", "min", "min", "max", "max")
stat_label <- c("Age Restrictions", "No Age Restrictions", "Age Restrictions", "No Age Restrictions", "Age Restrictions", "No Age Restrictions")
consWithWithoutAge <- data.frame(stat, stat_type, stat_label)

consWithWithoutAge %>%
  ggplot(aes(x = stat_label,
             y = stat,
             fill = factor(stat_label))) +
  geom_col() +
  facet_wrap(~stat_type) +
  xlab("") +
  ylab("Average Alcohol Consumption") +
  labs(title = "Average Alcohol Consumption in Countries\nWith and Without Drinking Age Restrictions", fill='Drinking Restrictions') +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_fill_brewer(palette = "Paired")
```

## Show Relevant Summary Statistic
```{r}
used <- c("Country", "Region", "Happiness Score", "GDP_PerCapita",
          "Beer_PerCapita", "Wine_PerCapita", "Spirit_PerCapita",
          "OnPremisePurchaseAge", "alcTotal")
summary(alcData[, used])
```

